Bayesian multivariate skew-normal finite mixture model for analysis of infant development trajectories

### 2019-06-24 manuscript/manuscript.tex
section 3

We can extend model (\ref{eq:msndens}) to the regression setting by modeling $\boldsymbol\zeta_{ki}$ as a function of covariates. Here we adopt a convenient stochastic representation of the MSN density (Azzalini and Dalla Valle, 1996):
\begin{equation}
\mathbf{y}_i|(z_i=k,t_i) = \mathbf{X}_i \boldsymbol\beta_k + t_i \boldsymbol\psi_k + \boldsymbol\epsilon_{ki}, \label{eq:msnreg}
\end{equation}
where 
$\mathbf{X}_i$ is a $J \times Jp$ design matrix that includes potential time-varying covariates (e.g., indicators denoting quarterly visits); $\boldsymbol{\beta}_k=(\beta_{k11},\ldots,\beta_{k1p},\ldots,\beta_{kJ1},\ldots,\beta_{kJp})^T$ is a $Jp\times 1$ vector of cluster- and outcome-specific regression coefficients; $t_i\sim N_{[0,\infty)}(0,1)$ is a subject-specific standard normal random variable truncated below by zero; $\boldsymbol\psi_k=(\psi_{k1},\ldots,\psi_{kJ})^T$ is a $J \times 1$ vector of cluster-specific skewness parameters; and $\boldsymbol{\epsilon}_{ki} \sim N_J(\boldsymbol0,\boldsymbol\Sigma_k)$ is a $J\times 1$ vector of error terms. Thus, conditional on $t_i$ and $z_i=k$, $\boldsymbol{y}_i$ is distributed as $N_J(\mathbf{X}_i \boldsymbol\beta_k + t_i \boldsymbol\psi_k, \boldsymbol{\Sigma}_k)$. Marginally (after integrating over $t_i$), $\mathbf{y}_i$ is distributed $MSN_J(\boldsymbol\zeta_{ik}, \boldsymbol\alpha_k, \boldsymbol\Omega_k)$, where through back-transformation \textbf{[Carter: Please review these expressions carefully and make sure they conform to those in FS and Pyne]}
\begin{eqnarray*}
\boldsymbol\zeta_{ik} &=& \mathbf{X}_i\boldsymbol\beta_k\\
\boldsymbol\alpha_k &=& \frac{1}{\sqrt{1 - \boldsymbol\psi_k^T \boldsymbol\Omega^{-1}_k\boldsymbol\psi_k}} \boldsymbol\Omega^{-1}_k\boldsymbol\psi_k~~\text{and}\\
\boldsymbol\Omega_k &=& \boldsymbol\Sigma_k  + \boldsymbol\psi_k \boldsymbol\psi_k^T.
\end{eqnarray*}
Additional details can be found in Fr\:{u}wirth-Schnatter and Pyne (2010).

>>>>

We can extend model (\ref{eq:msndens}) to the regression setting by modeling $\boldsymbol\zeta_{ki}$ as a function of covariates. Here we adopt a convenient stochastic representation of the MSN density (Azzalini and Dalla Valle, 1996):
\begin{equation}
\mathbf{y}_{i}|(z_i=k,t_i) = \mathbf{X}_i \boldsymbol\beta_k + t_i \boldsymbol\psi_k + \boldsymbol\epsilon_{ki}, \label{eq:msnreg}
\end{equation}
where 
$\mathbf{X}_i$ is a $J \times Jp$ design matrix that includes potential time-varying covariates (e.g., indicators denoting quarterly visits); $\boldsymbol{\beta}_k=(\beta_{k11},\ldots,\beta_{k1p},\ldots,\beta_{kJ1},\ldots,\beta_{kJp})^T$ is a $Jp\times 1$ vector of cluster- and outcome-specific regression coefficients; $t_i\sim N_{[0,\infty)}(0,1)$ is a subject-specific standard normal random variable truncated below by zero; $\boldsymbol\psi_k=(\psi_{k1},\ldots,\psi_{kJ})^T$ is a $J \times 1$ vector of cluster-specific skewness parameters; and $\boldsymbol{\epsilon}_{ki} \sim N_J(\boldsymbol0,\boldsymbol\Sigma_k)$ is a $J\times 1$ vector of error terms. Thus, conditional on $t_i$ and $z_i=k$, $\boldsymbol{y}_i$ is distributed as $N_J(\mathbf{X}_i \boldsymbol\beta_k + t_i \boldsymbol\psi_k, \boldsymbol{\Sigma}_k)$. Marginally (after integrating over $t_i$), $\mathbf{y}_i$ is distributed $MSN_J(\boldsymbol\zeta_{ki}, \boldsymbol\alpha_k, \boldsymbol\Omega_k)$, where through back-transformation 
\begin{eqnarray*}
\boldsymbol\zeta_{ki} &=& \mathbf{X}_i\boldsymbol\beta_k\\
\boldsymbol\alpha_k &=& \frac{1}{\sqrt{1 - \boldsymbol\psi_k^T 
\boldsymbol\Omega^{-1}_k\boldsymbol\psi_k}} \boldsymbol\omega_k \boldsymbol\Omega^{-1}_k\boldsymbol\psi_k~~\text{and}\\
\boldsymbol\Omega_k &=& \boldsymbol\Sigma_k  + \boldsymbol\psi_k \boldsymbol\psi_k^T,
\end{eqnarray*}
where $\boldsymbol\omega_k = \text{Diag}(\sqrt{\omega_{k,11}},...,\sqrt{\omega_{k,JJ}})$ is the $J \times J$ diagonal matrix containing the square root of the diagonal entries of $\boldsymbol\Omega_k$. Additional details can be found in Fr\"{u}wirth-Schnatter and Pyne (2010).
